---
title: "Data Visualization Pre-Lecture"
author: "Dr. Osita Onyejekwe"
date: "2023-01-30"
output: html_document
---

-   **Data**
-   **Mapping (aesthetics)**
-   **geometric representation**
-   statistics
-   facet
-   coordinate space
-   labels
-   theme

```{r}

#install.packages("tidyverse")
#install.packages("scales")install.packages("ggplot2", dependencies = TRUE)
#install.packages("ggplot2", dependencies = TRUE)

#library(tidyverse)
library(ggplot2)

# type in data() to see the list of available data sets in R. For this lecture we are going to be looking at the BOD (Biochemical Oxygen Demand) data set. 

# type in ?BOD to see the description of this data set. 


```

Let's take a look at the data:

# Method 1

```{r}
#install.packages("colorspace")
#library(colorspace)


library(ggplot2)
ggplot(data = BOD, mapping = aes(x = Time, y= demand)) + 
  geom_point(size  = 2 ) + geom_line(color = "red")
```

# Method2

```{r}

ggplot(BOD, aes(Time, demand)) +
  geom_point(size=2) + geom_line(color = "blue")

```

## Carbon Dioxide Uptake in Grass Plants

```{r}

library(tidyverse)




# Becuase we are working in the Tidyverse, we have access to the pipe operator. 


# conerntration is the CO2 concerntration in the atmosphere (independent)
# uptake of C02 (y variable), dependent 

CO2 %>% ggplot(aes(conc, uptake, color = Treatment)) +
  geom_point(size = 2, alpha = 0.5) + geom_smooth() 
# alpha is for transparency  
# geom_smooth() is a linear model that will map over our geom_point()
# as things get plotted, I want the color of individaul plots to be a function of which treatment group that particular observation or row was int 

# It's drawn 2 smooth linear models through these 2 groups, the chilled and non-chilled (treatment groups applied to graphs) and around each of them it has drawn the standard error. 

 
CO2 %>% ggplot(aes(conc, uptake, color = Treatment)) +
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm, se =F)


# Create a Facet wrap by type. 

  
CO2 %>% ggplot(aes(conc, uptake, color = Treatment)) +
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm, se =F)+
 facet_wrap(~Type)

CO2 %>% ggplot(aes(conc, uptake, color = Treatment)) +
  geom_point(size = 2, alpha = 0.5) + geom_smooth(method = lm, se =F)+
 facet_wrap(~Type) + labs(title = "Concentration of C02") + theme_bw()


```

```{r}

CO2 %>% ggplot(aes(Treatment, uptake)) + geom_boxplot()
  
CO2 %>% ggplot(aes(Treatment, uptake)) + geom_boxplot() + geom_point()
# now let's add in mapping just for the points 

CO2 %>% ggplot(aes(Treatment, uptake)) + geom_boxplot() + geom_point(alpha = 0.5 , aes(size = conc, color = Plant))

# here note that for geom_point(aes(....)) the mapping here only applied to the points. It did not apply to the other layers of this particular graph. 


# let's now flip it on its side. 

CO2 %>% ggplot(aes(Treatment, uptake)) + geom_boxplot() + geom_point(alpha = 0.5 , aes(size = conc, color = Plant)) + facet_wrap(~Type) + coord_flip() + theme_bw() +labs(title = "Chilled vs Non-Chilled")
# face_wrap(~Type) to get more information out of it. 

```

## In Summary

We've got a number of different attributes being represented here. The type of plant, Quebec/Mississippi, the actual name/designation of the plant (bottom right panel), concentration, uptake, we've got treatment and a lot of attributes all summarized in one graphic.

```{r}


#view(mpg)

# we may postulate that the engine size (displ) may translate to better "city" and "hwy" fuel efficienty (miles per gallon) 

mpg %>% ggplot(aes(displ, cty)) + geom_point()

# We observe that as engine size grows bigger, fuel efficiency goes down. 


mpg %>% ggplot(aes(displ, cty)) + geom_point(aes(color = drv, size = trans, alpha = 0.5))

# aesthetic where we map color against the drive (4 wheel/front wheel..etc) such that the size of each dot is a function of the transmission type. 

mpg %>% ggplot(aes(displ, cty)) + geom_point(aes(color = drv, size = trans, alpha = 0.5)) + geom_smooth(method = lm)



mpg %>% ggplot(aes(displ, cty)) + geom_point(aes(color = drv, size = trans, alpha = 0.5)) + geom_smooth(method = lm) + facet_wrap(~year, nrow = 1 )

```

Let's say that we wanted to clean this graph up a little bit by removing some outliars such that the city data is filtered by only observations less than 25.

```{r}

mpg %>% filter(cty <25) %>% ggplot(aes(displ, cty)) + geom_point(aes(color = drv, size = trans, alpha = 0.5)) + geom_smooth(method=lm) + facet_wrap(~year, nrow = 1) + labs(x = "Engine Size", title = "Fuel Efficiency") + theme_bw()
```
